---
title: "Statistical inference course project"
output:
  pdf_document: default
  html_document: default
Author: Sara Moujahid
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

Part 1: Simulation Exercise Instructions :

In this project you will investigate the exponential distribution in R and compare it with the Central Limit Theorem. The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter.
Set lambda = 0.2 for all of the simulations. You will investigate the distribution of averages of 40 exponentials. Note that you will need to do a thousand simulations.

Simulating the exponential distribution :
```{r Exponential Distribution}
set.seed(16)
lambda <-0.2
n <- 40
sim <- 1000
#We create 1K times a sample of 40 exponentials, then compute its mean
exp_dis_sample <- replicate (sim, mean(rexp (n,lambda)))
```

Calculating the sample mean :
```{r Sample mean}
mean(exp_dis_sample)
```
Calculating the theoretical mean :
```{r Theoretical mean}
1/lambda
```
```{r Comparison means}
##There is a very small difference between the theoretical and sample means of this exponential dsitribution
abs(mean(exp_dis_sample)-1/lambda)
```
Calculating the sample variance :
```{r Sample var}
var(exp_dis_sample)
```
Calculating the theoretical variance for a population of n:
```{r theoretical var}
(1/lambda)^2/n
```

```{r Comparison vars}
##There is a very small difference between the theoretical and sample variances of this exponential dsitribution
abs(var(exp_dis_sample)-(1/lambda)^2/n)
```

Plotting the distribution, and comparing it to a normal distribution of similar mean and standard deviation :
```{r Histogram distribution, echo=FALSE}
#The historgram is roughly symmetric and mound shaped, like a normal distribution due to CLT. The more we increase the sample size, the histogram will resemble more the normal distribution
hist(exp_dis_sample,prob=TRUE, breaks=40,main='Histogram of avg of 40 Exponentials for 1K simulations',col='red')
x<-seq(min(exp_dis_sample),max(exp_dis_sample),length=2*n)
y<-dnorm(x,1/lambda,1/(lambda*sqrt(n)))
lines(x,y , pch=22, col='black', lwd=2)

```

Part 2: Basic Inferential Data Analysis Instructions :

Now in the second portion of the project, we're going to analyze the ToothGrowth data in the R datasets package.

```{r tootgrowth}
library(datasets)
df<-data(ToothGrowth)
#First rows of the data sets
head(ToothGrowth)
#length of the dataset
dim(ToothGrowth)
#Some statistics summaries of the dataset's columns
summary(ToothGrowth)
# Display different summary
str(ToothGrowth)
#Only 3 values for dose
unique(ToothGrowth$dose)

#Some plots to better understand the data :
#Transform dose into factors
ToothGrowth$dose<-as.factor(ToothGrowth$dose)
#contruct a boxplot
library(ggplot2)
ggplot(data = ToothGrowth, aes(x = dose, y = len)) + 
  geom_boxplot(aes(fill = dose), width = 0.8) + xlab('Dose') + ylab('Tooth length') + facet_grid(~supp) +ggtitle('Tooth Length vs. Doses, by supp')
```

Calculating confidence intervals :

Hypothesis1 : The supplements offer the same tooth growth
```{r hyp1}

# Run a t-test on the data
t.test(len~supp,data=ToothGrowth)
```
the p-value is 6%, so higher than 5%, and the confidence interval contains 0. The null hypothesis cannot be rejected.


Hypothesis2 : The supplements offer the same tooth growth For the dosage of '0.5mg/day'
```{r hyp2}
# Run a t-test on the data for dose .5mg/day
df<-subset(ToothGrowth,dose==.5)
t.test(len~supp,data=df)
```
the p-value is .6%, so less than 5%, and the confidence interval is higher than 0. So the results are statistically significant and the null hypothesis is rejected. The OJ supp offer more tooth growth than VC.

```{r hyp3}
# Run a t-test on the data for dose 1mg/day
df<-subset(ToothGrowth,dose==1)
t.test(len~supp,data=df)
```
the same is true for a dosage of 1mg/day

```{r hyp4}
# Run a t-test on the data for dose 2mg/day
df<-subset(ToothGrowth,dose==2)
t.test(len~supp,data=df)
```
Unfortunately for a dosage of 2mg/day, the p-value is very high, and the confidence interval contains 0. So the null hypothesis cannot be rejected.

Conclusion :
* For the dosages of .5 and 1mg/day, the OJ supplement offers more tooth growth than VC. However, we failed to reject the null hypothesis for the dosage 2mg/day and conclude there is no significant difference between both supplements.